---
title: "Splines"
output: 
  html_document:
    toc: true
    toc_float: true
---

## Context and Data

Before proceeding, install the `splines` package by entering `install.packages("splines")` in the Console.

We'll continue using the `College` dataset in the `ISLR` package to explore splines. You can use `?College` in the Console to look at the data codebook.

```{r}
library(caret)
library(ggplot2)
library(dplyr)
library(readr)
library(ISLR)
library(splines)

data(College)

# A little data cleaning
college_clean <- College %>% 
    mutate(school = rownames(College)) %>% 
    filter(Grad.Rate <= 100) # Remove one school with grad rate of 118%
rownames(college_clean) <- NULL # Remove school names as row names
```


## Exercise 1: Evaluating a fully linear model

We will model `Grad.Rate` as a function of 4 predictors: `Private`, `Terminal`, `Expend`, and `S.F.Ratio`.

a. Make scatterplots with 2 different smoothing lines to explore potential nonlinearity. Adding the following to the normal scatterplot code will create a smooth (curved) blue trend line and a red linear trend line.

```{r}
geom_smooth(color = "blue", se = FALSE) +
geom_smooth(method = "lm", color = "red", se = FALSE)
```

b. Use `caret` to fit an ordinary linear regression model (no splines yet) with the following specifications:

- Use 8-fold CV.
- Use mean absolute error (MAE) to select a final model.
- Select the simplest model for which the metric is within one standard error of the best metric.

```{r}
set.seed(___)
ls_mod <- train(

)
```

c. Make plots of the residuals vs. the 3 quantitative predictors to evaluate the appropriateness of linear terms.

```{r}
ls_mod_data <- college_clean %>%
    mutate(
        pred = predict(ls_mod, newdata = college_clean),
        resid = ___
    )

ggplot(ls_mod_data, ???) +
    ___ +
    ___ +
    geom_hline(yintercept = 0, color = "red")
```





## Exercise 2: Evaluating a spline model

We'll extend our linear regression model with spline functions of the quantitative predictors (leave `Private` as is).

a. What tuning parameter is associated with splines? How do high/low values of this parameter relate to bias and variance?

b. Update your code from Exercise 1 to model the 3 quantitative predictors with natural splines that have 2 knots (= 3 degrees of freedom).

```{r}
set.seed(___)
spline_mod <- train(

)
```

c. Make plots of the residuals vs. the 3 quantitative predictors to evaluate if splines improved the model.

```{r}
spline_mod_data <- ___

# Residual plots

```

## Extra! Variable scaling

What is your intuition about whether variable scaling matters for the performance of splines?

Check you intuition by reusing code from Exercise 2, except with `preProcess = "scale"` inside `train()`. Call this `spline_mod2`.

How do the predictions from `spline_mod` and `spline_mod2` compare? You could use a plot to compare or check out the `all.equal()` function.

